---
title: "Fitting a MSFA model via the ECM Algorithm"
author: "Roberta de Vito, Ruggero Bellio"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting a MSFA model via the ECM Algorithm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The following example illustrates how to fit a MSFA model via the ECM Algorithm, 
using a data set available in the Bioconductor repository (www.bioconductor.org). 

## Getting the data
Some pre-processing is required to get the data into a form suitable for the
analysis. This was already done, and the resulting data frame is saved into the
`data_immune` object. The commands that were used to form it are included
in the help file for the data object.

```{r help, echo = TRUE, results = TRUE, tidy = TRUE}
library(MSFA)
data(data_immune)
help(data_immune)
```

## Obtaining suitable starting values for model parameters
Then we get suitable starting values for model parameters, selecting $K=3$ common
factors and $3, 4$ study-specific factors for the two studies, respectively.

```{r, starting values, messages = FALSE}
start_value <- start_msfa(X_s = data_immune, k = 3, j_s = c(3, 4))
```

## Fitting the model via ECM
Now everything is in place for estimating the model parameters via the ECM algorithm

```{r, get estimate, results = FALSE}
mle <-  ecm_msfa(data_immune, start_value, trace = FALSE)
```

The estimated matrix of common loadings can be represented by a suitable heatmap:

```{r, heatmap, fig.show = 'hold', fig.width = 7.5, fig.height = 6.5, message = FALSE}
library(gplots)
heatmap.2(mle$Phi,dendrogram='none', Rowv=FALSE, Colv=FALSE,trace='none', density.info="none", col=heat.colors(256))
```
